---
title: "Lab 8"
author: "Margaret Connor"
date: "3/13/2019"
output: html_document
---
**Objective: Learn how to apply Neural network using R.**

In the field of engineering, it is crucial to have accurate estimates of the performance of building materials. These estimates are required in order to develop safety guidelines governing the materials used in the construction of buildings, bridges, and roadways.

Estimating the strength of concrete is a challenge of particular interest. Although it is used in nearly every construction project, concrete performance varies greatly due to a wide variety of ingredients that interact in complex ways.

As a result, it is difficult to accurately predict the strength of the final product. A model that could reliably predict concrete strength given a listing of the composition of the input materials could result in safer construction practices.

# Collecting data
For this analysis, we will utilize data on the compressive strength of concrete donated to the UCI Machine Learning Data Repository by I-Cheng Yeh.

As he found success using neural networks to model these data, we will attempt to replicate his work using a simple neural network model in R.

According to the website, the concrete dataset contains 1,030 examples of concrete with eight features describing the components used in the mixture.

These features are thought to be related to the final compressive strength and they include the amount (in kilograms per cubic meter) of cement, slag, ash, water, superplasticizer, coarse aggregate, and fine aggregate used in the product in addition to the aging time (measured in days).

# Exploring and preparing the data
**1. Load the data for analysis.**
```{r}
concrete <-read.csv('/Users/margaretconnor/NetBeansProjects/Rdata/concrete.csv')
```

**2. Look into the data structure and describe the features you found.**
```{r}
str(concrete)
```

**3. How to prepare the data before applying ANN? Test your updates.**
```{r}
```


**4. Partition the data into a training set with 75 percent of the examples, concrete_train, and a testing set with 25 percent, concrete_test.**

